Very simplified likelihood

### 2023-11-28 inst/very_simplified_likelihood.Rmd
bespoke statistical test

```{r maxvscheco}
a <- read.table("formula1_2023.txt",header=TRUE)
a <- a[,seq_len(ncol(a)-1)]
max <- as.numeric(as.matrix(a)[1,])
checo <- as.numeric(as.matrix(a)[2,])
checo[is.na(checo)] <- 22
max
checo
MLE(max,22)
MLE(checo,22)
```

>>>>

```{r maxvscheco}
a <- read.table("formula1_2023.txt",header=TRUE)
a <- a[,seq_len(ncol(a)-1)]
d1 <- as.numeric(as.matrix(a)["Ocon",])
d2 <- as.numeric(as.matrix(a)["Gasly",])
d1[is.na(d1)] <- 22
d2[is.na(d2)] <- 22
d1
d2
MLE(d1,22)
MLE(d2,22)
```

two sample test:

```{r 2samptest}
both <- c(d1,d2)
bothmax <- MLE(both,22)$maximum

d1max <- MLE(d1,22)$maximum
d2max <- MLE(d2,22)$maximum
f(bothmax,d1,22,log=TRUE)
f(bothmax,d2,22,log=TRUE)
f(d1max,d1,22,log=TRUE)
f(d2max,d2,22,log=TRUE)

Lambda <- (f(d2max,d1,22,log=TRUE)+f(d2max,d2,22,log=TRUE)) -(f(bothmax,d1,22,log=TRUE)+f(bothmax,d2,22,log=TRUE))
Lambda
pchisq(-2*Lambda,df=1,lower.tail=FALSE)
```

## Mathematical Olympiad

```{r readoly}
sametest <- function(d1,d2,n){
  both <- c(d1,d2)
  bothmax <- MLE(both,n)$maximum

  d1max <- MLE(d1,n)$maximum
  d2max <- MLE(d2,n)$maximum
  Lambda <- (f(d2max,d1,n,log=TRUE)+f(d2max,d2,n,log=TRUE)) -(f(bothmax,d1,n,log=TRUE)+f(bothmax,d2,n,log=TRUE))
  pchisq(-2*Lambda,df=1,lower.tail=FALSE)
}

a <- as.matrix(read.table("olympiad.txt"))
nrow(a)
d1 <- a["AUS",]
d2 <- a["NZL",]
d1
d2
sametest(d1,d2,100)
a["USA",]
a["CHN",]
sametest(a["USA",],a["CHN",],100)
sametest(a["USA",],a["CHN",],5)
```
